EGGs are adhesive!

### 2025-07-28 jlamp/jlamp.tex
revisione

\newcommand{\lgh}{\mathsf{lg}}
\newcommand{\eq}{\mathsf{eq}}
\newcommand{\quo}{\mathsf{Q}}

@@

\def\R{\mathsf{R}}
\def\B{\textbf {\textup{B}}}
\def\C{\textbf {\textup{C}}}
\def\D{\textbf {\textup{D}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\def\E{\textbf {\textup{E}}}
\def\T{\textbf {\textup{1}}}
\def\A{\textbf {\textup{A}}}
\def\M{\mathcal{M}}
%categorie varie
\newcommand{\catname}[1]{\textbf{\textup{#1}}}
\newcommand{\lab}{\catname{LHyp}}
\newcommand{\hyp}{\catname{Hyp}}
\newcommand{\hyps}{\catname{Hyp}_{\Sigma}}
\newcommand{\EqHyp}{\catname{EqHyp}} %equivalence hypergraphs
\newcommand{\EqHyps}{\catname{EqHyp}_{\Sigma}}
\newcommand{\EqTG}{\catname{EqTG}}
\newcommand{\EqTGs}{\catname{EqTG}_{\Sigma}}
\newcommand{\GEqTGs}{\catname{GEqTG}_{\Sigma}}
\newcommand{\GEqATGs}{\catname{GEqATG}_{\Sigma}}
\newcommand{\gr}{\textbf{\textup{Graph}}}
\newcommand{\dgr}{\catname{SGraph}}
\newcommand{\dg}{\catname{DAG}}
\newcommand{\rt}{\mathsf{dcl_s}}
\newcommand{\rta}{\mathsf{dcl}}
\newcommand{\rtd}{\mathsf{dcl_{d}}}
\newcommand{\slice}[2]{(\catname{#1}\downarrow{#2})}
\newcommand{\tg}[0]{\catname{TG}_{\Sigma}}
\newcommand{\teg}[0]{\catname{TeGr}_{\Sigma}}
\newcommand{\sv}[0]{\mathsf{Sieves}}
\newcommand{\mono}[1]{\mathsf{Mon}(\catname{#1})}
\newcommand{\mo}[1]{{#1}_\mathsf{Mon}}
\newcommand{\pro}{\mathsf{prod}}
\newcommand{\spro}{\mathsf{ps}}
\newcommand{\prol}{\mathsf{lprod}}
\newcommand{\pred}[1]{{\downarrow}#1}
\newcommand{\colim}[0]{\mathrm{colim}}
\newcommand{\cod}{\mathsf{cod}}
\renewcommand{\sp}{\mathsf{sp}}
\renewcommand{\sup}{\mathsf{sup}}
\newcommand{\cow}[1]{\mathsf{cwd}({#1})}
\renewcommand{\inf}{\mathsf{inf}}
\newcommand{\dom}{\mathsf{dom}}
\newcommand{\dwnarrow}{\downarrow \hspace{-2pt}}
\newcommand{\Dwnarrow}{\Downarrow \hspace{-2pt}}
\newcommand{\egg}{\mathsf{e}\text{-}\catname{EqHyp}}

\newcommand{\ari}{\mathsf{ar}}
\newcommand{\abs}[1]{\lvert #1\rvert}

@@

@@

\emph{E-graphs} (shortly, EGGs) are an up-and-coming formalism for program optimisation and synthesis via a compact 
representation and efficient implementation of equality saturation. 
%
Albeit a classical data structure~\cite{DetlefsNS05}, EGGs received 
new impulse after the seminal~\cite{WillseyNWFTP21} and
developed a thriving community, as witnessed by the official website~\cite{eggs}.
%
The key idea of rewriting-based program optimisation is to perform the manipulation of a syntactical description 
of a program, replacing some of its components in such a way that the semantics is preserved while 
the computational costs of its actual execution are improved. Instead of directly removing sub-programs, EGGs just add the 
new components and link them to the older ones via the equivalence relation, until an optimal program is 
reached and extracted.

@@

%
%\begin{notation}
We open this background section by fixing some notation.
%
Given a category $\X$ we do not distinguish notationally between $\X$ and its class of objects, so
``$X\in \X$'' means that $X$ is an object of $\X$. We let $\mor(\X)$, $\mon(\X)$ and $\reg(\X)$ denote the class of all arrows, monos and regular monos of $\X$, respectively.  Given an object $X$, we  denote by $?_X$ the unique arrow from an initial object into $X$ and by $!_X$ that  unique arrow from $X$ into a terminal one. We will also use the notation $e\colon X\eto Y$ to denote that an arrow $e\colon X\to Y$ is a regular epi. 
%\end{notation}

@@

\noindent
\begin{minipage}[l]{.85\linewidth}
	\begin{itemize}
		\item		\emph{stable under pushouts (pullbacks)} if for every pushout (pullback) square as the one aside, if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
		\item \emph{closed under composition} if $h, k\in \mathcal{A}$ implies $h\circ k\in \mathcal{A}$ whenever $h$ and $k$ are composable;
		\item \emph{closed under decomposition}, or \emph{left-cancellable}, if whenever $g$ and $g\circ f$ belong to $\mathcal{A}$, then $f\in \mathcal{A}$.
	\end{itemize}
\end{minipage}\hfill 
\begin{minipage}[r]{.13\linewidth}
\xymatrix{A \ar[r]^{f} \ar[d]_{m}& B \ar[d]^{n} \\ C \ar[r]_{g} & D}
\end{minipage}

@@

\noindent 
\begin{minipage}[l]{.85\linewidth}
\begin{defi} A \emph{kernel pair} for an arrow $f\colon A \to B$ is an object $K_f$ together with two arrows $\pi^1_f, \pi^2_f\colon K_f \rightrightarrows A$, denoted as $(K_f, \pi^1_f, \pi^2_f)$, such that the square aside is a pullback.
\end{defi}
\end{minipage} \hfill 
\begin{minipage}[r]{.18\linewidth}
	\xymatrix@R=15pt{K_f \ar[r]_{\pi^2_f} \ar[d]_{\pi^1_f}& A \ar[d]^f \\ A \ar[r]^{f}& B}
\end{minipage}

@@

\noindent
\begin{minipage}[l]{.6\linewidth}
\begin{restatable}{prop}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top face is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}
\end{minipage}\hfill 
\begin{minipage}[r]{.38\linewidth}\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }
	\end{minipage}

@@

Focusing on $\Set$, we can prove a sharper result (see \Cref{proof:kerset} for the proof).
\smallskip

@@

\begin{prop}\label{prop:fact}
	Let $X$ be a set and $n\in \mathbb{N}$. Then the following facts hold
	\begin{enumerate}
		\item there are arrows $v_{n}\colon X^n\to X^\star $ such that $(X^\star, \{v_{n}\}_{n\in \mathbb{N}})$ is a coproduct;
		\item for every arrow $f\colon X\to Y$, $f^\star\colon X^\star \to Y^\star$ is the coproduct of the family $\{f^n\}_{n\in \mathbb{N}}$;
		\item $(-)^\star$ preserves all \emph{connected limits} \cite{carboni1995connected}, in particular it preserves pullbacks and equalizers.
	\end{enumerate}
\end{prop}

@@

\begin{rem}\label[remark]{rem:length}Notice that $1^\star$ can be canonically identified with $\mathbb{N}$, thus for every set $X$ the arrow $!_\X\colon X\to 1$ induces a \emph{length function} $\lgh_{X}\colon X^\star \to \mathbb{N}$, which sends a word to its length.
\end{rem}

@@

We open this section with the definition of hypergraphs and we show how to label them with an algebraic signature.

@@

We define $\hyp$ to be the resulting category.
\end{defi}\end{minipage} \hfill 
\begin{minipage}[r]{.15\linewidth}
\xymatrix@R=8pt{E_{\mathcal{G}} \ar[d]_{h} \ar[r]_{s_{\mathcal{G}}}& V^\star_{\mathcal{G}}  \ar[d]^{k^\star}\\E_{\mathcal{G}} \ar[r]^{s_{\mathcal{H}}} & V^\star_{\mathcal{H}}} \hspace{1pt}\\ \xymatrix@R=8pt{E_{\mathcal{G}} \ar[d]_{h} \ar[r]_{t_{\mathcal{G}}}& V^\star_{\mathcal{G}}  \ar[d]^{k^\star}\\E_{\mathcal{G}} \ar[r]^{t_{\mathcal{H}}} & V^\star_{\mathcal{H}}} 
\end{minipage}

@@

\begin{prop}\label[proposition]{prop:com}
	$\hyp$ is isomorphic to $\comma{\id{\Set}}{\pro^\star}$
\end{prop}

Note that by hypothesis $(-)^\star$ preserves pullbacks, while $\pro$ is continuous by definition, 
hence by \Cref{prop:com} and \Cref{cor:mono} we can deduce the following result.

@@

\begin{exa}\label[example]{exa_2}
We represent hypergraphs visually: dots denote nodes and boxes hyperedges. Should we be interested in their identity, we put a name near the corresponding dot or box. 
Sources and targets are represented by lines between dots and squares: the lines from the sources of a hyperedge comes from the left of the box, while the lines to the targets 
exit from the right of the box.
Let us look at the picture below. It represent a hypergraph $\mathcal{G}$ with sets $V_\mathcal{G} = \{v_1, v_2, v_3, v_4\}$ and $E_\mathcal{G} = \{h_1, h_2, h_3, h_4\}$ of nodes 
and edges, respectively, such that  $h_1, h_2$ have no source and $h_3, h_4$ a pair of nodes
	\begin{center}
        \begin{tikzpicture}
                \begin{pgfonlayer}{nodelayer}
                        \node[style=none] (lab1) at (-1.8, 0.8) {$h_1$};
                        \node[style=small box] (h1) at (-1.8, 0.3) {};
                        \node[style=small box] (2) at (-1.8, -0.5) {};
                        \node[style=none] (lab2) at (-1.8, -1.0) {$h_2$};
                        \node[style=none] (h1b) at (-1.8, 0.3) {};
                        \node[style=none] (th2) at (0, 0.3) {};
                        \node[style=none] (bh2) at (0, -0.3) {};
                        \node[style=none] (h2b) at (0, 0) {};
                        \node[style=none] (th3) at (1.8, 0){};
                        \node[style=none] (bh3) at (1.8, -0.6){};
                        \node[style=none] (h3c) at (1.8, -0.3){};
                        \node[style=none] (2c) at (-1.8, -0.5){};
                        \node[style=none] (lab3) at (0, 0.8) {$h_3$};
                        \node[style=medium box] (h2) at (0, 0) {};
                        \node[style=none] (lab4) at (1.8, 0.5) {$h_4$};
                        \node[style=medium box] (h3) at (1.8, -0.3) {};
                        \node[style=none](labv2) at (-0.9, -0.9) {$v_2$};
                        \node[style=node] (v2) at (-0.9, -0.5) {};
                        \node[style=none] (labv1) at (-0.9, 0.7) {$v_1$};
                        \node[style=node] (v1) at (-0.9, 0.3) {};
                        \node[style=none] (labv3) at (0.9, 0.4) {$v_3$};
                        \node[style=node] (v3) at (0.9, 0) {};
                        \node[style=none] (labv4) at (2.7, 0.1) {$v_4$};
                        \node[style=node] (v4) at (2.7, -0.3) {};
                \begin{pgfonlayer}{edgelayer}
                        \draw (h1b.center) to (th2.center);
                        \draw (2c.center) to (v2);
                        \draw[in=180, out=30] (v2) to (bh2.center);
                        \draw (h2b.center) to (th3.center);
                        \draw (v4) to (h3c.center);
                        %\draw (h3c) to (v4);
                        \draw[in=180, out=-30](v2) to (bh3);
	\end{center}
%We  decorate the arrow corresponding to the $i^{th}$ letter of a target or a source with a label $i$.
%\todo{Ho lasciato un unico esempio (che è anche un term graph): occorre riscrivere questo paragrafo una volta che abbiamo deciso come rappresentare i grafi}
\iffalse 
\begin{exa}
Take $V_{\mathcal{G}}$ to be be $\{v_1, v_2, v_3, v_4, v_5\}$ and $E_{\mathcal{G}}$ to be $\{h_1, h_2, h_3\}$. Sources and targets are given by:
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}}  & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4 
		\end{matrix} && s_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_5\\
		t_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4
		\end{matrix} && t_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & 0\mapsto v_5 && t_{\mathcal{G}}(h_3)\colon 0\to V_{\mathcal{G}} &  t_{\mathcal{G}}(h_3)=?_{ V_{\mathcal{G}}} 
	\end{matrix}\]

	We can draw the resulting $\mathcal{G}$ as follows:
	\begin{center}\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_1$}] (A) at (0,0) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_2$}] (B) at (0,-1.5) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_3$}] (C) at (3,0) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_4$}] (D) at (3,-1.5) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_5$}] (E) at (6,-0.75) {};
			\draw[rounded corners] (1.25, -1) rectangle (1.75, -0.5) {};
			\draw[->-=.5](4.75,-0.75)--(E)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5](E)--(7,-0.75)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (4.25, -1) rectangle (4.75, -0.5) {};
			\node at (4.5, -0.3){$h_2$};
			\node at (1.5, -0.3){$h_1$};
			\node at (7.25, -0.3){$h_3$};
			\draw[rounded corners] (7, -1) rectangle (7.5, -0.5) {};
			\draw(A)[->-=.5]..controls(0.5,0)and(1.2,-0.2)..(1.25,-0.6)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw(B)[->-=.5]..controls(0.5,-1.5)and(1.2,-1.3)..(1.25,-0.9)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};

			\draw(C)[->-=.5]..controls(3.5,0)and(4.2,-0.25)..(4.25,-0.6)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw(D)[->-=.5]..controls(3.5,-1.5)and(4.2,-1.3)..(4.25,-0.9)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};

			\draw[->-=.5](1.75,-0.9)..controls(1.8,-1.3)and(2.5,-1.5)..(D)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5] (1.75,-0.6)..controls(1.8,-0.25)and(2.5,0)..(C) node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
		\end{tikzpicture}
	\end{center}
\end{exa}

\begin{exa}\label[example]{exa_21} Let $V_{\mathcal{G}}$ be as in the previous example and $E_{\mathcal{G}}=\{h_1, h_2, h_3\}$.	Then we define
	\[\begin{matrix}
		A && s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}}  & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix}&& s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
			0 \mapsto v_1\\
			1\mapsto v_4	
		\end{matrix}\\
			B && t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_1 && 		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&   t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5
	\end{matrix}\]

>>>>

\newcommand{\lgh}{\mathsf{lg}}
\newcommand{\ind}[1]{\mathsf{in}_{\mathcal{#1}}}
\newcommand{\ino}[1]{\mathsf{IN}({\mathcal{#1}})}
\newcommand{\outd}[1]{\mathsf{out}_{\mathcal{#1}}}
\newcommand{\eq}{\mathsf{eq}}
\newcommand{\quo}{\mathsf{Q}}

@@

\def\R{\mathsf{R}}
\def\B{\textbf {\textup{B}}}
\def\C{\textbf {\textup{C}}}
\def\D{\textbf {\textup{D}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\def\E{\textbf {\textup{E}}}
\def\T{\textbf {\textup{1}}}
\def\A{\textbf {\textup{A}}}
\def\M{\mathcal{M}}
%categorie varie
\newcommand{\catname}[1]{\textbf{\textup{#1}}}
\newcommand{\lab}{\catname{LHyp}}
\newcommand{\lmo}{\catname{LMH}}
\newcommand{\hyp}{\catname{Hyp}}
\newcommand{\shyp}{\catname{SepHyp}}
\newcommand{\hyps}{\catname{Hyp}_{\Sigma}}
\newcommand{\EqHyp}{\catname{EqHyp}} %equivalence hypergraphs
\newcommand{\EqHyps}{\catname{EqHyp}_{\Sigma}}
\newcommand{\EqTG}{\catname{EqTG}}
\newcommand{\EqTGs}{\catname{EqTG}_{\Sigma}}
\newcommand{\GEqTGs}{\catname{GEqTG}_{\Sigma}}
\newcommand{\GEqATGs}{\catname{GEqATG}_{\Sigma}}
\newcommand{\gr}{\textbf{\textup{Graph}}}
\newcommand{\dgr}{\catname{SGraph}}
\newcommand{\dg}{\catname{DAG}}
\newcommand{\rt}{\mathsf{dcl_s}}
\newcommand{\rta}{\mathsf{dcl}}
\newcommand{\rtd}{\mathsf{dcl_{d}}}
\newcommand{\slice}[2]{(\catname{#1}\downarrow{#2})}
\newcommand{\tg}[0]{\catname{TG}_{\Sigma}}
\newcommand{\teg}[0]{\catname{TeGr}_{\Sigma}}
\newcommand{\sv}[0]{\mathsf{Sieves}}
\newcommand{\mono}[1]{\mathsf{Mon}(\catname{#1})}
\newcommand{\mo}[1]{{#1}_\mathsf{Mon}}
\newcommand{\pro}{\mathsf{prod}}
\newcommand{\spro}{\mathsf{ps}}
\newcommand{\prol}{\mathsf{lprod}}
\newcommand{\pred}[1]{{\downarrow}#1}
\newcommand{\colim}[0]{\mathrm{colim}}
\newcommand{\cod}{\mathsf{cod}}
\renewcommand{\sp}{\mathsf{sp}}
\newcommand{\is}[1]{{#1}_{is}}
\renewcommand{\sup}{\mathsf{sup}}
\newcommand{\cow}[1]{\mathsf{cwd}({#1})}
\renewcommand{\inf}{\mathsf{inf}}
\newcommand{\dom}{\mathsf{dom}}
\newcommand{\dwnarrow}{\downarrow \hspace{-2pt}}
\newcommand{\Dwnarrow}{\Downarrow \hspace{-2pt}}
\newcommand{\egg}{\mathsf{e}\text{-}\catname{EqHyp}}

\newcommand{\ari}{\mathsf{ar}}
\newcommand{\coar}{\mathsf{co}}
\newcommand{\abs}[1]{\lvert #1\rvert}

@@

%questo sparirà, serve a me per non diventare scemo
\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}
\tableofcontents

@@

\emph{E-graphs} (shortly, EGGs) are an up-and-coming formalism for program optimisation and synthesis via a compact 
representation and efficient implementation of equality saturation. Albeit a classical data structure \cite{DetlefsNS05}, EGGs received 
new impulse after the seminal \cite{WillseyNWFTP21} and
developed a thriving community, as witnessed by the official website \cite{eggs}.
%
The key idea of rewriting-based program optimisation is to perform the manipulation of a syntactical description 
of a program, replacing some of its components in such a way that the semantics is preserved while 
the computational costs of its actual execution are improved. Instead of directly removing sub-programs, EGGs just add the 
new components and link them to the older ones via the equivalence relation, until an optimal program is 
reached and extracted.

@@

We open this background section by fixing some notation. Given a category $\X$ we do not distinguish notationally between $\X$ and its class of objects, so
``$X\in \X$'' means that $X$ is an object of $\X$. We let $\mor(\X)$, $\mon(\X)$ and $\reg(\X)$ denote the class of all arrows, monos and regular monos of $\X$, respectively.  Given an object $X$, we  denote by $?_X$ the unique arrow from an initial object into $X$ and by $!_X$ that  unique arrow from $X$ into a terminal one. We will also use the notation $e\colon X\eto Y$ to denote that an arrow $e\colon X\to Y$ is a regular epi.

@@

\noindent
\begin{minipage}[l]{.85\linewidth}
	\begin{itemize}
		\item		\emph{stable under pushouts (pullbacks)} if for every pushout (pullback) square as the one aside, if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
		\item \emph{closed under composition} if $h, k\in \mathcal{A}$ implies $h\circ k\in \mathcal{A}$ whenever $h$ and $k$ are composable;
	\end{itemize}
\end{minipage}\hfill 
\begin{minipage}[r]{.13\linewidth}
\xymatrix{A \ar[r]^{f} \ar[d]_{m}& B \ar[d]^{n} \\ C \ar[r]_{g} & D}
\end{minipage}
\begin{itemize}	\item \emph{closed under decomposition}, or \emph{left-cancellable}, if whenever $g$ and $g\circ f$ belong to $\mathcal{A}$, then $f\in \mathcal{A}$.
\end{itemize}

@@

\noindent 
\begin{minipage}[l]{.85\linewidth}
\begin{defi} A \emph{kernel pair} for an arrow $f\colon A \to B$ is an object $K_f$ together with two arrows $\pi^1_f, \pi^2_f\colon K_f \rightrightarrows A$, denoted as $(K_f, \pi^1_f, \pi^2_f)$, such that the square aside is a pullback.
\end{defi}
\end{minipage} \hfill 
\begin{minipage}[r]{.18\linewidth}
	\xymatrix@R=15pt{K_f \ar[r]^{\pi^2_f} \ar[d]_{\pi^1_f}& A \ar[d]^f \\ A \ar[r]^{f}& B}
\end{minipage}

@@

\noindent
\begin{minipage}[l]{.6\linewidth}
\begin{restatable}{prop}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top face is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}
\end{minipage}\hfill 
\begin{minipage}[r]{.38\linewidth}\xymatrix@C=10pt@R=2pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]_(.3){f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]_{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]_(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]_{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]^{g} & & D }
	\end{minipage}

@@

Focusing on $\Set$, we can prove a sharper result (see \Cref{proof:kerset} for the proof).
\medskip

@@

\begin{prop}\label[prop]{prop:fact}
	Let $X$ be a set and $n\in \mathbb{N}$. Then the following facts hold
	\begin{enumerate}
		\item there are arrows $v_{n}\colon X^n\to X^\star $ such that $(X^\star, \{v_{n}\}_{n\in \mathbb{N}})$ is a coproduct;
		\item for every arrow $f\colon X\to Y$, $f^\star\colon X^\star \to Y^\star$ is the coproduct of the family $\{f^n\}_{n\in \mathbb{N}}$;
		\item $(-)^\star$ preserves all \emph{connected limits} \cite{carboni1995connected}, in particular it preserves pullbacks and equalizers.
	\end{enumerate}
\end{prop}

@@

\noindent
\begin{minipage}[l]{.83\linewidth}
	\begin{rem}\label[remark]{rem:length} It is well known that $(\mathbb{N}), \{\delta_n\}_{n\in \mathbb{N}})$ is a coproduct for $\aleph_0$ copies of $1$, where $\delta_n\colon 1\to \mathbb{N}$ is the function picking the element $n$. Thus by the first point of \Cref{prop:fact} there is a canonical isomorphism $\Delta\colon \mathcal{N}\to 1^\star$. In particular, every set $X$ the arrow $!_\X\colon X\to 1$ induces a \emph{length function} $\lgh_{X}\colon X^\star \to \mathbb{N}$, which sends a word to its length, as in the diagram aside.
	\end{rem}
\end{minipage}\hfill \begin{minipage}[r]{.15\linewidth}\noindent
\xymatrix@R=10pt{1 \ar[r]^{\id{1}} \ar[d]^{u_n}& 1 \ar[d]^{v_n}\\ \mathbb{N} \ar@{.>}[r]^{\Delta}& 1^\star} \hspace{1pt}\\
\xymatrix@R=10pt@C=22pt{\mathbb{N} \ar[r]^{\Delta} \ar@{.>}[dr]_{\lgh_{X}}& 1^\star\\ & X^\star \ar[u]_{!^\star_X}}
\end{minipage}

\begin{notaz}
	Given a word $w\in X^\star$ of length $l$, for every $1\leq i \leq l$, we will denote by $w_i$ the $i^\mathrm{th}$ letter of $w$.  
\end{notaz}

@@

Let us begin this section with the definition of hypergraphs.

@@

We define $\hyp$ to be the resulting category.
\end{defi}\end{minipage} \hfill 
\begin{minipage}[r]{.15\linewidth}
\xymatrix@R=8pt{E_{\mathcal{G}} \ar[d]_{h} \ar[r]^{s_{\mathcal{G}}}& V^\star_{\mathcal{G}}  \ar[d]^{k^\star}\\E_{\mathcal{G}} \ar[r]^{s_{\mathcal{H}}} & V^\star_{\mathcal{H}}} \hspace{1pt}\\ \xymatrix@R=8pt{E_{\mathcal{G}} \ar[d]_{h} \ar[r]^{t_{\mathcal{G}}}& V^\star_{\mathcal{G}}  \ar[d]^{k^\star}\\E_{\mathcal{G}} \ar[r]^{t_{\mathcal{H}}} & V^\star_{\mathcal{H}}} 
\end{minipage}

@@

\begin{prop}[prop]\label{prop:com}
	$\hyp$ is isomorphic to $\comma{\id{\Set}}{\pro^\star}$
\end{prop}

Note that by, \Cref{prop:fact}, $(-)^\star$ preserves pullbacks, while $\pro$ is continuous by definition, 
hence by \Cref{prop:com} and \Cref{cor:mono} we can deduce the following result.

@@

\begin{exa}\label[example]{exa_2}
We represent hypergraphs visually: dots denote nodes and boxes hyperedges. Should we be interested in their identity, we put a name near the corresponding dot or box. 
Sources and targets are represented by lines between dots and squares: the lines from the sources of a hyperedge comes from the left of the box, while the lines to the targets 
exit from the right of the box.
Let us look at the picture below. It represent a hypergraph $\mathcal{G}$ with sets $V_\mathcal{G} = \{v_1, v_2, v_3, v_4\}$ and $E_\mathcal{G} = \{h_1, h_2, h_3, h_4\}$ of nodes 
and edges, respectively, such that  $h_1, h_2$ have no source and $h_3, h_4$ a pair of nodes
	\begin{center}
        \begin{tikzpicture}
                \begin{pgfonlayer}{nodelayer}
                        \node[style=none] (lab1) at (-1.8, 0.8) {$h_1$};
                        \node[style=small box] (h1) at (-1.8, 0.3) {};
                        \node[style=small box] (2) at (-1.8, -0.5) {};
                        \node[style=none] (lab2) at (-1.8, -1.0) {$h_2$};
                        \node[style=none] (h1b) at (-1.8, 0.3) {};
                        \node[style=none] (th2) at (0, 0.3) {};
                        \node[style=none] (bh2) at (0, -0.3) {};
                        \node[style=none] (h2b) at (0, 0) {};
                        \node[style=none] (th3) at (1.8, 0){};
                        \node[style=none] (bh3) at (1.8, -0.6){};
                        \node[style=none] (h3c) at (1.8, -0.3){};
                        \node[style=none] (2c) at (-1.8, -0.5){};
                        \node[style=none] (lab3) at (0, 0.8) {$h_3$};
                        \node[style=medium box] (h2) at (0, 0) {};
                        \node[style=none] (lab4) at (1.8, 0.5) {$h_4$};
                        \node[style=medium box] (h3) at (1.8, -0.3) {};
                        \node[style=none](labv2) at (-0.9, -0.9) {$v_2$};
                        \node[style=node] (v2) at (-0.9, -0.5) {};
                        \node[style=none] (labv1) at (-0.9, 0.7) {$v_1$};
                        \node[style=node] (v1) at (-0.9, 0.3) {};
                        \node[style=none] (labv3) at (0.9, 0.4) {$v_3$};
                        \node[style=node] (v3) at (0.9, 0) {};
                        \node[style=none] (labv4) at (2.7, 0.1) {$v_4$};
                        \node[style=node] (v4) at (2.7, -0.3) {};
                \end{pgfonlayer}
                \begin{pgfonlayer}{edgelayer}
                        \draw (h1b.center) to (th2.center);
                        \draw (2c.center) to (v2);
                        \draw[in=180, out=30] (v2) to (bh2.center);
                        \draw (h2b.center) to (th3.center);
                        \draw (v4) to (h3c.center);
                        %\draw (h3c) to (v4);
                        \draw[in=180, out=-30](v2) to (bh3);
                \end{pgfonlayer}
        \end{tikzpicture}
	\end{center}
\end{exa}

A rich source of example is given by \emph{monoidal signatures}.

\smallskip 
\noindent
\begin{minipage}[l]{.75\linewidth}
	\begin{defi}
		A \emph{(single typed) monoidal signature} $\Sigma$ is a triple $(O_\Sigma, \ari_\Sigma, \coar_\Sigma)$ given by a \emph{set of operations} $O_\Sigma$ together with \emph{arity} and \emph{coarity} functions $\ari_\Sigma, \coar_\Sigma\colon O_\Sigma \rightrightarrows \mathbb{N}$.

		We say that $\Sigma$ is \emph{algebraic} if $\coar_{\Sigma}$ is the constant function in $1$.

		We define the \emph{hypergraph $\mathcal{G}_\Sigma$ associated with $\Sigma$} taking $O_\Sigma$ as set of hyperedges, $1$ as set of nodes and $\Delta \circ \ari_\Sigma$ as the source function and $ \Delta \coar_{\Sigma}$ as target function, where $\Delta \colon \mathbb{N}\to 1$ is the isomorphism of \Cref{rem:length}.
	\end{defi}

	\begin{exa}\label{exa_3} 
		Given a set $A$, let $\Sigma = (O_\Sigma, \ari_\Sigma, \coar_{\Sigma})$ be the algebraic signature with $O_\Sigma =  \mathbb{N} + A + \{*, /\}$, 
		and 
		\[\ari_\Sigma(n)=0 \quad \ari_\Sigma(a) = 0, \quad \ari_\Sigma(*) =2 \quad  \ari_\Sigma(/) = 2\]
		for every $n\in \mathbb{N}$ and $a\in A$. Then we can depict the hypergraph $\mathcal{G}^\Sigma$  as in the picture aside.
	\end{exa}
	\end{minipage}
	%\begin{center}
	\hfill
	\begin{minipage}[r]{.22\linewidth}
	\begin{tikzpicture}
		\begin{pgfonlayer}{nodelayer}
			\node[style=node] (v) at (0, 0){};
			\node[style=none] (vlab) at (0, -0.4){$v$};
